---
title: "Wildcards in the Bash Shell"
description: |
  CSC405 Project Online Tutorial. This tutorial will discuss some basic wildcard     commands to use when you are using a large amount of files at one time.
author:
  - name: Reagan Atchison
    url: 
date: 04-17-2019
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
library(mdsr)
library(tidyverse)
library(ggplot2)
library(stringr)
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this blog post, I have created a tutorial on how to perform basic wildcard commands in bash. I will be using a directory that includes subdirectories, and within those subdirectories there are files with random names. The file types include .R and .txt. I will be using wildcards in order to show how to find files that include certain characters, how to replace a character in file names to a different character, and more.

Bash, also known as the "Bourne Again Shell" is a shell that allows a human to interact with a computer. Bash allows the user to communicate with the computer through command promtps and go through the computer's memory to find and edit files. Included within bash, there are wildcards that allow a user to search and edit many different things at one time. In this tutorial, I will be showing basic ways to use these wildcards.  

To begin, you will need to run this code in order to create the directory used throughout this tutorial:

```{r}
library(stringr)
if (dir.exists("myDir")) {
  system(paste0("rm -rf myDir"))
}

fruit2 <-
  fruit %>% 
  str_replace_all(pattern = " ", replacement = "_")

dir.create("mydir")
for (i in 1:5) {
  dirName <- paste0("mydir/example", i)
  dir.create(dirName)
  for (letter in letters) {
    file.create(paste0(dirName, "/", "file_", letter, ".txt"))
  }
  for (col in colors()) {
    file.create(paste0(dirName, "/", col, "_", sample(fruit2, size = 1), ".R"))
  }
}
```

Within this directory that has been created, five subdirectories will be created along with a lot of random files within them. This is the only code that you will need to run to follow along with this tutorial, yet results may vary because the files created are named at random. 

After you run the R script, you should see a new directory that has been created in your home directory called `mydir`. This is the directory that we will be using. The first command in all chunks will be changing the directory, and then the code to complete the process will be next.

## Tutorial

### The Asterisk Wildcard

One way you can search for different file names within a large directory is by using the * wildcard. The * wildcard allows a user to specify an amount of characters a filename should include, and the * allows the rest of the characters to be random. The example shown below allows a user to list all of the file names that start with "al" and can have any characters following it.

```{bash}
cd ~/mydir/example1
ls al*
```

There is only one output from this specific directory, which is shown above. This means that there were no other instances when a file name started with "al."

### The Question Mark Wildcard

Let's say that you wanted to look at a group of files under the same name, but maybe they all had a different, single character on the ending. In order to do this, you would not want to use the * wildcard. Instead, you would use the ? wildcard. While the * wildcard allows a file to have an unlimited number of characters after the *, the ? only searches for files that have one character following the ? wildcard. This is helpful in cases where you have different versions of a file, and you want to see all of the versions. I have shown an example of this below.

```{bash}
cd ~/mydir/example1
ls file_?.txt
```

In this example, the output shows us all the files that begin with "file_" and end with ".txt." The ? wildcard is allowing that single character to be anything, and if the rest of the specified text is included in the same order, then it will be included in the output.

### The Brackets Wildcards

From the files shown above, maybe you only want to see the files that end with a letter that is in the first half of the alphabet. 

In order to do this, you will need to use the [ ] (brackets) wildcard, with the letters you would like to see within the brackets. I have shown an example below:

```{bash}
cd ~/mydir/example1
ls file_[a-m].txt
```

By using this wildcard, we are able to see all files that end in a letter in the first half of the alphabet.

You can also use brackets if you wanted to find files that have a range of numbers in them. In the example below, I wish to see all of the files that begin with "cadetblue" and have any of the numbers 1-3 following it, along with any characters that come after that:

```{bash}
cd ~/mydir/example2
ls cadetblue[1-3]*
```

The use of theses wildcards together let the computer know that I only wish to see the files that begin with "cadetblue" followed by one number in the range of 1-3, along with any characters that come after. The only specifications that I made in this command were that the file had to start with "cadetblue" and the number that comes after that must be within the range 1-3.

Instead, maybe you would like to see all of files that start with "cadetblue" that do not have the number 1-3 following it. You can look at that too, by simply placing an exclamation point (!) before the first character in the range. I have shown an example below:

```{bash}
cd ~/mydir/example2
ls cadetblue[!1-3]*
```

The output shows us results in which there is not a 1-3 listed after the color. We see a 4 and we see no number, but we do not see any numbers within the range of 1-3.

You can also do this with character that are not in sequence with each other. For example, maybe you only want to see the files that start with "file_" and end with either an a, c, x, or z. We can do that with brackets, as shown below:

```{bash}
cd ~/mydir/example1
ls file_[acxz].txt
```

The output shows us the files that end in an a, c, x, or z. We can do this with capital letter aswell, but they would need to be included in the brackets as a separate letter.

### The `cat` Wildcard

If you would like to see the contents of a file, you can do that with the `cat` wildcard. For example, if I were to type "hello" in my file titled "aliceblue_satsuma.R", I could print that out by using `cat`. I have shown that example below:

```{bash}
cd ~/mydir/example1
cat aliceblue_satsuma.R
```

Since there is only one line, and one word in this file, that is all it prints. I typed "hello" in this file, and the `cat` wildcard returns the input. If you would like, you can try this on your own machine by typing something into a file, and use the `cat` command followed by the filename to return what you typed. 

### Renaming Multiple Files at One Time

While I was not able to do this with a wildcard, I thought it would be important to include this in my tutorial. If you would like to change a lot of files to have a different name, this can be done with a for-loop. I have shown an example below of how a for-loop is structured in bash:

```{bash}
cd ~/mydir/example3

for i in turquoise*
do
    mv "$i" "${i/turquoise/umbrella}"
done
```

This for loop looks very similar to the for-loops in R. You have to tell bash to  "do" whatever you want it to complete within the for-loop, and you have to tell bash when it can stop working by saying "done." "mv" means to move the characters, in this case it is searching for the word "turquoise" in a filename, and then replacing it with the word "umbrella." If we go through the file list, we can see that it worked and all the filenames with "turquoise" in it got changed to "umbrella." If you wanted to change what files it looks for and what it changes it to, you would simply change those two words in the for-loop, within the "mv" command. 
Citation:(https://stackoverflow.com/questions/20657432/rename-multiple-files-but-only-rename-part-of-the-filename-in-bash/20657563)



## Conclusion

In conclusion, wildcards play a very important role while using bash. Wildcards make using bash a lot easier, and much more simple. There are many more wildcards other than the ones that I have discussed here, but I have included most of the common ones.

## References

Newham, Cameron, and Bill Rosenblatt. Learning the Bash Shell. OReilly, 2005.

https://stackoverflow.com/questions/20657432/rename-multiple-files-but-only-rename-part-of-the-filename-in-bash/20657563
